*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-29 Code/Manuscript_O2.Rmd
Naaman and Sylwia edits added

Metaproteomic datasets were obtained from the KM1128 entry in the BCO-DMO database (https://www.bco-dmo.org/deployment/59053) accessed via the OPP in June 2019. Datasets contained:
	i) Peptide sequences and sample identification (ID) number;
	ii) Sample ID number, station, depth in meters below the surface the sample was collected at, best-hit BLASTP protein and species annotation and the corresponding Uniprot Entry number for the identified proteins;
	iii) Sample station depth and [O~2~].  
	The depth and [O~2~] were joined to peptide sequence and BLASTP annotations by ID number, depth and station using tidyverse package [@R-tidyverse] running under R vXXX and RStudio v1.2.5019 (https://posit.co/)  (LOCATION OF CODE ON GITHUB). The resulting merged dataset was filtered for those *Prochlorococcus* peptides, detected from 0 to 300 m below the surface, annotated as a subunit of *Prochlorococcus* chlorophyll binding proteins (Pcb); Photosystem II (PSII); Cytochrome b~6~f (Cytb~6~f);  Photosystem I (PSI); NADPH Dehydrogenase (NDH); Plastoquinol Terminal Oxidase (PTOX); Plastocyanin (PC); Ferredoxin (Fd); Ribulose-1,5-bisphosphate oxygenase (RUBISCO);  ATP Synthase;  FtsH proteases (FtsH) or ribosomes. 
	Detected peptides were re-annotated for consistency and labelled, where feasible, according to strain, clade, subunit and protein complex. Full protein sequences corresponding to detected proteins were obtained from UniProt (https://www.uniprot.org/) and analyzed in Molecular Evolution and Genetic Analyses X (MEGAX) software  (https://www.megasoftware.net/). Sequences for proteins for each of the thirteen Prochlorococcus strains identified in the dataset were aligned with MUSCLE using UPGMA cluster method and a lambda of 24 with a -2.9 gap open penalty and 1.20 hydrophobicity multiplier.  Overall mean pairwise distance between protein sequences was determined using bootstrap variance estimation methods. Maximum likelihood phylogenetic trees were assembled using 1000 bootstrap replications with a 95% site coverage cut off. *Prochlorococcus* FtsH isoform identities, and functions, were inferred by sequence comparisons to the characterized four isoforms of FtsH protease of *Synechocystis* sp. PCC6803 (CITATONS).  Data for each strain was plotted against depth and  [O~2~] and sampling station.

@@

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA
Maine, US (REFERENCE). MED4 (CCMP1986) is from High-Light adapted (HLI) clade; SS120
(CCMP1375) is from Low-Light adapted (LLIII) clade; and MIT9313 (CCMP2773)
is from Low-Light adapted (LLIV) clade. Cultures were
maintained in incubators set to 22°C with a light/dark cycle of 12 h. The PAR level for maintenance cultures reflected Photosynthetically Active Radiation (PAR) in the source
niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^ with illumination from STANDARD Products Inc. Cool White F24T5/41K/8/HO/PS/G5/STD, 24 watts, fluorescent bulbs; SS120 and
MIT9313 at 30 µmol photons m^-2^ s^-1^ with illumination from Philips Cool White F14T5/841 Alto, 14 watts, fluorescent bulbs. To maintain active growth all strains were
transferred weekly with 1 in 5 dilutions with Pro99 media
[@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial
seawater [@bergesEvolutionArtificialSeawater2001].

Controlled growth experiments were performed using MCMIX-OD or MC1000-OD PSI Multicultivators (Figure \@ref(fig:LabeledMC); PSI, Drásov, Czech Republic).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (White LED, 660 nm, 450 nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~)
gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system (Qubit Systems Inc, Kingston, Ontario, Canada). [O~2~] in situ was verified using oxygen optodes (PyroScience,
Germany) inserted into tubes for
real-time measurements, with a temperature probe in
the aquarium of the bioreactor to correct [O~2~] measures for temperature
fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the
salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05%
CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

*In situ* measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the
duration of the lag phase, if any. All data from the Multicultivator were saved
as comma separated values files  (https://github.com/FundyPhytoPhys/prochlorococcus_o2/tree/main/Data/RawData).

@@

## Growth rate analysis {.unnumbered}
Data files (.csv) saved from the Multicultivator software were imported into R-Studio [@R-Studio] for data management [@R-tidyverse], growth rate calculations CITATION, comparisons of model fits [@harrisonPhotosynthesisirradianceRelationshipsPolar1986], and generation of
figures using `ggplot2` [@R-ggplot2]. The chlorophyll proxy
optical density (OD680 - OD720;  ΔOD) was used to determine the chlorophyll specific growth
rate (µ, d^-1^) for each treatment combination.  We first used a rolling mean from the RStudio `zoo` package [@R-zoo] to calculate the average ΔOD data over a 1-hour window to lower influence of outlier points and remove data points collected during post stationary phase, if applicable. We used the Levenberg-Marquardt algorithm
[@bellaviaLevenbergMarquardtMethod2018] modification of the non-linear
least squares,  using the R package `minpack.lm`
[@R-minpack.lm], to fit a logistic
equation \@ref(eq:GrowthRate)

```{=tex}
\begin{equation}
  µ = \frac{ΔOD_{max} × ΔOD_{min} × exp^{(µ × t)}}{ΔOD_{max} + (ΔOD_{min} × exp^{((µ × t) - 1)})}
  (\#eq:GrowthRate)
\end{equation}
```

@@

A Generalized Additive Model (GAM) [@woodGeneralizedAdditiveModels2017] was applied to the
relation between chlorophyll-specific µ, d^-1^  to photoperiod and PAR level,  for each growth [O~2~] level, and for the blue and red wavebands for growth, for each *P.marinus* strain in this study. The gam function from the R package `mgcv` [@R-mgcv] was used to model the growth rate with smoothing terms and indicate the 90, 50 and 10% quantiles for growth rate across the levels of factors. Only growth rate estimates for which the amplitude of standard error was smaller than 30% of the fitted growth rate were included in the GAM modelling. Our priority was the effects of ecologically relevant blue light on growth trends.  We included analyses of growth responses to red light, which is not ecophysiologically relevant, but which might prove mechanistically informative [@murphyPhotoinactivationPhotosystemII2017].

@@

To estimate the Photosynthetic Usable Radiation (PUR), a proxy of incident photons absorbed by the cells, for each *P. marinus* ecotype, the imposed Photosynthetic Active
Radiation (PAR) was first determined using the reported delivery of sinusoidal diel PAR regimes by the Multicultivators, point validated using an XXXXXcitation of PAR meterXXX. An emission profile from 400 nm to 700 nm of each coloured LED
light of the MCMIX-OD Multicultivator and the white LED light of the MC1000-OD Multicultivator was obtained using a Jaz spectrometer (Ocean Optics, Inc.,Dunedin, FL, USA) equipped with a fiber
optic cable, HH2 FiberOpticJmp (Part number A901073, Malvern Panalytical
Ltd, Malvern, UK). Each LED spectrum was then normalized to its emission maximum.
An *in-vivo* whole cell absorbance spectrum for each *P. marinus* strain under each spectral growth condition was obtained using the Olis 14 UV/VIS Clarity
Spectrophotometer (Olis Inc, Bogart, GA, USA) to scan across range of λ = 350 nm to 750 nm at 1 nm intervals. The path length of the internally reflective cavity of the Olis spectrophotometer was corrected to a 1 cm path length using the Javorfi correction method [@javorfiQuantitativeSpectrophotometryUsing2006] on PRO 99 media subtracted whole cell absorbance spectra. The blank-corrected whole cell absorbance spectra were normalized to the absorbance maximum of divinyl chlorophyll *a* (Chl a~2~), determined for each spectra, falling between 400 nm and 460 nm.

An integrated weighting Equation \@ref(eq:WeightedPUR) [@morelAvailableUsableStored1978] was used to determine the weighted PUR spectrum P(λ),

@@

Protein samples were thawed and placed in 2 mL FastPrep®-24 bead lysing matrix D vials (MP Biomedicals, Irvine, CA, USA; SKU 116913050) with 200 uL of 1X protein extraction buffer (0.0281 mol L^−1^ TRIS base buffer, 0.0211 mol L^−1^ TRIS HCl buffer, 0.0735 mol L^−1^ lithium dodecyl sulfate, 1.08 mol L^−1^ glycerol, 0.5 mmol L^−1^ ethylenediaminetetraacetic acid, 0.1 mg·mL−1 PefaBloc SC (AEBSF) protease inhibitor) [@brownFluxCapacitiesAcclimation2008]  XXX note we used a low tris PEB recipe not sure where it came from (citation??) XXX. Cells were lysed for 3 cycles of 60 seconds at 6.5m s-1 with 60 second intervals on ice between cycles.  The lysing vials were centrifuged at 16100g using XXXXXX centrifuge ( reference here    ) for 5 minutes, the supernatant was transferred to a 600 uL microcentrifuge tube and centrifuged again at 16100g for 3 minutes. 40 µL of the supernatant were subsampled into labeled vials and stored at -80 °C until analysis. The bicinchoninic acid (BCA) assay was used for total protein determination using a calibration curve of bovine gamma globulin (BGG; Bio-Rad, Hercules, CA, USA) standard ranging from 0 to 1.0 mg mL^-1^ range.

@@

Using the dataset Omar *et al*. [@omarAnnotationGenesEncoding2023], we filtered the dataset for Enzyme Commission Numbers (EC numbers), or Kegg Orthology Numbers (KO numbers) identified by BRENDA  [@changBRENDAELIXIRCore2021] as 'natural substrates' for O~2~ in *P. marinus* (MED4, MIT9313, SS120, and NATL2A) in order to generate a subset containing only those orthologs encoding enzymes directly mediating O~2~ metabolism. We grouped orthologs together by EC number and their KO number (KO number) and determined the occurrences of individual orthologs encoding each EC number, or KO number when EC number was not available, in a given strain. We merged the dataset with a list of enzyme Michaelis constant (K~m~) values from other organisms, as K~m~ values from *Prochlorococcus* were only available in the case of Ribulose bisphosphate carboxylase. XXX Possibly add the link to the full list of genes. XXXX

@@

NATLProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

Guided by the evidence of ocean distributions of proteins from *Prochlorococcus* we set up a matrix of photoperiods, PAR, spectral bands, and [O~2~] to approximate current, and potential future, latitudinal, depth and seasonal niches for *Prochlorococcus* strains.  We earlier found [@murphyPhotoinactivationPhotosystemII2017] that growth under red light could prove mechanistically informative as to factors limiting *Prochlorococcus* growth, so we included that spectral band even though it is not characteristic of representative *Prochlorococcus* niches.  Although *Prochlorococcus* is currently limited to a narrow range of surface photoperiods, potential poleward latitudinal expansions, in combination with attenuation of light with depth, mean *Prochlorococcus* may potentially encounter a wide range of photoperiods.  Our growth rate determinations generally agreed with those from @mooreComparativePhysiologySynechococcus1995, for white LED and 250 µM O~2~, but our study is, we believe, the first to analyze the interactive growth responses of *Prochlorococcus* strains to varying [O~2~], spectral wavebands and photoperiods.

@@

Similar to growth trends under 250 µM O~2~, MED4 maintained at 25 µM O~2~ showed fastest growth when the
photoperiod was 16 h for each spectral waveband, across PAR levels (Figure \@ref(fig:MED4PAR)).
The µ~max~ was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 180 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod. The 4 h photoperiod
experiments under white LED light were not performed because no growth was achieved when grown under an 8 h photoperiod of white LED light.

MED4 did not grow under when sparged to the lowest [O~2~] of 2.5 µM (Figure \@ref(fig:MED4PAR)). 2.5 µM O~2~ growth experiments
were not conducted for 4 and 16 h photoperiods, as no reproducible growth occurred when MED4 was exposed to 8 and 12 h photoperiods.

@@

Under 250 µM O~2~ growth rates of *P. marinus* SS120, Clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white, red or blue LED light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

Under 25 µM O~2~ and PAR of 30 µmol photons m^-2^ s^-1^ growth trends were similar to 250 µM O~2~. SS120 showed no
growth under a 4 h photoperiod for red spectral waveband, however under blue light, SS120 was able to grow under a 4 h photoperiod (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew under 4 and 8 h blue light and 16 h red light photoperiods, however the growth rate decreased under 12 and 16 h white LED light photoperiod treatments. Blue light treatments under PAR of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h photoperiod. The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ blue light and 8 h photoperiod. The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^  under white LED light experiments were not performed due to time constraints.

@@

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments. Most 4 h photoperiod treatments of SS120 also did not grow under 250 µM O~2~, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments(Figure \@ref(fig:BluevsRedPurFitsPlots)B).

>>>>

Metaproteomic datasets were obtained from the KM1128 entry in the BCO-DMO database (https://www.bco-dmo.org/deployment/59053) accessed via the OPP in June 2019. Datasets contained:
	i) Peptide sequences and sample identification (ID) number;
	ii) Sample ID number, station, depth in meters below the surface the sample was collected at, best-hit BLASTP protein and species annotation and the corresponding Uniprot Entry number for the identified proteins;
	iii) Sample station depth and [O~2~].  
	The depth and [O~2~] were joined to peptide sequence and BLASTP annotations by ID number, depth and station using *tidyverse* package [@R-tidyverse] running under R vXXX and RStudio v1.2.5019 (https://posit.co/)  (LOCATION OF CODE ON GITHUB). The resulting merged dataset was filtered for those *Prochlorococcus* peptides, detected from 0 to 300 m below the surface, annotated as a subunit of *Prochlorococcus* chlorophyll binding proteins (Pcb); Photosystem II (PSII); Cytochrome b~6~f (Cytb~6~f);  Photosystem I (PSI); NADPH Dehydrogenase (NDH); Plastoquinol Terminal Oxidase (PTOX); Plastocyanin (PC); Ferredoxin (Fd); Ribulose-1,5-bisphosphate oxygenase (RUBISCO);  ATP Synthase;  FtsH proteases (FtsH) or ribosomes. 
	Detected peptides were re-annotated for consistency and labelled, where feasible, according to strain, clade, subunit and protein complex. Full protein sequences corresponding to detected proteins were obtained from UniProt (https://www.uniprot.org/) and analyzed in Molecular Evolution and Genetic Analyses X (MEGAX) software  (https://www.megasoftware.net/). Sequences for proteins for each of the thirteen Prochlorococcus strains identified in the dataset were aligned with MUSCLE using UPGMA cluster method and a lambda of 24 with a -2.9 gap open penalty and 1.20 hydrophobicity multiplier.  Overall mean pairwise distance between protein sequences was determined using bootstrap variance estimation methods. Maximum likelihood phylogenetic trees were assembled using 1000 bootstrap replications with a 95% site coverage cut off. *Prochlorococcus* FtsH isoform identities, and functions, were inferred by sequence comparisons to the characterized four isoforms of FtsH protease of *Synechocystis* sp. PCC6803 (CITATONS).  Data for each strain was plotted against depth and  [O~2~] and sampling station.

@@

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA
Maine, US (REFERENCE). MED4 (CCMP1986) is from High-Light adapted (HLI) clade; SS120
(CCMP1375) is from Low-Light adapted (LLIII) clade; and MIT9313 (CCMP2773)
is from Low-Light adapted (LLIV) clade. Cultures were
maintained in incubators set to 22°C with a light/dark cycle of 12 h. The PAR level for maintenance cultures reflected PAR)in the source niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^ with illumination from STANDARD Products Inc. Cool White F24T5/41K/8/HO/PS/G5/STD, 24 watts, fluorescent bulbs; SS120 and
MIT9313 at 30 µmol photons m^-2^ s^-1^ with illumination from Philips Cool White F14T5/841 Alto, 14 watts, fluorescent bulbs. To maintain active growth all strains were
transferred weekly with 1 in 5 dilutions with Pro99 media
[@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial
seawater [@bergesEvolutionArtificialSeawater2001].

Controlled growth experiments were performed using MCMIX-OD or MC1000-OD PSI Multicultivators (Figure \@ref(fig:LabeledMC); PSI, Drásov, Czech Republic).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (White LED, 660 nm, 450 nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~)
gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system (Qubit Systems Inc., Kingston, Ontario, Canada). [O~2~] in situ was verified using oxygen optodes (PyroScience,
Germany) inserted into tubes for
real-time measurements, with a temperature probe in
the aquarium of the bioreactor to correct [O~2~] measures for temperature
fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the
salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05%
CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

*In situ* measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the
duration of the lag phase, if any. All data obtained from the Multicultivator were saved
as comma separated values files  (https://github.com/FundyPhytoPhys/prochlorococcus_o2/tree/main/Data/RawData).

@@

## Growth rate analysis {.unnumbered}
Data files (.csv) saved from the Multicultivator software were imported into R-Studio for data management [@R-tidyverse], growth rate calculations CITATION, comparisons of model fits [@harrisonPhotosynthesisirradianceRelationshipsPolar1986], and visualization. The chlorophyll proxy
optical density (OD~680~ - OD~720~;  ΔOD) was used to determine the chlorophyll specific growth
rate (µ, d^-1^) for each treatment combination.  We first used a rolling mean from the R package *`zoo`* [@R-zoo] to calculate the average ΔOD data over a 1-hour window to lower the influence of outlier points and remove data points collected during post stationary phase, when applicable. We used the Levenberg-Marquardt algorithm
[@bellaviaLevenbergMarquardtMethod2018] modification of the non-linear
least squares,  using the R package *`minpack.lm`*
[@R-minpack.lm], to fit a logistic
equation (Equation \@ref(eq:GrowthRate))

\begin{equation}
  µ = \frac{ΔOD_{max} × ΔOD_{min} × exp^{(µ × t)}}{ΔOD_{max} + (ΔOD_{min} × exp^{((µ × t) - 1)})}
  (\#eq:GrowthRate)
\end{equation}

@@

A Generalized Additive Model (GAM) [@woodGeneralizedAdditiveModels2017] was applied to the
relation between chlorophyll-specific µ, d^-1^  to photoperiod and PAR level,  for each growth [O~2~] level, and for the blue and red wavebands for growth, for each *P.marinus* strain in this study. The R package *`mgcv`* [@R-mgcv] was used to model the growth rate with smoothing terms and indicate the 90, 50 and 10% quantiles for growth rate across the levels of factors. Only growth rate estimates for which the amplitude of standard error was smaller than 30% of the fitted growth rate were included in the GAM. Our priority was the effects of ecologically relevant blue light on growth trends.  We included analyses of growth responses to red light, which is not ecophysiologically relevant, but which might prove mechanistically informative [@murphyPhotoinactivationPhotosystemII2017].

@@

To estimate the Photosynthetic Usable Radiation (PUR), a proxy of incident photons absorbed by the cells, for each *P. marinus* ecotype, the imposed Photosynthetic Active
Radiation (PAR) was first determined using the reported delivery of sinusoidal diel PAR regimes by the Multicultivators, point validated using a LI-250 quantum sensor (LI-COR Inc.,Lincoln, NE, USA). An emission profile from 400 nm to 700 nm of each coloured LED
light of the MCMIX-OD Multicultivator and the white LED light of the MC1000-OD Multicultivator was obtained using a Jaz spectrometer (Ocean Optics, Inc.,Dunedin, FL, USA) equipped with a fiber
optic cable, HH2 FiberOpticJmp (Part number A901073, Malvern Panalytical
Ltd, Malvern, UK). Each LED spectrum was then normalized to its emission maximum.
An *in-vivo* whole cell absorbance spectrum for each *P. marinus* strain under each spectral growth condition was obtained using the Olis 14 UV/VIS Clarity
Spectrophotometer (Olis Inc., Bogart, GA, USA) to scan across range of λ = 350 nm to 750 nm at 1 nm intervals. The path length of the internally reflective cavity of the Olis spectrophotometer was corrected to a 1 cm path length using the Javorfi correction method [@javorfiQuantitativeSpectrophotometryUsing2006] on PRO 99 media subtracted whole cell absorbance spectra. The blank-corrected whole cell absorbance spectra were normalized to the absorbance maximum of divinyl chlorophyll *a* (Chl a~2~), determined for each spectra, falling between 400 nm and 460 nm.

An integrated weighting equation \@ref(eq:WeightedPUR) [@morelAvailableUsableStored1978] was used to determine the weighted PUR spectrum P(λ),

@@

Protein samples were thawed and placed in 2 mL FastPrep®-24 bead lysing matrix D vials (MP Biomedicals, Irvine, CA, USA; SKU 116913050) with 200 µL of 1X protein extraction buffer (0.0281 mol L^−1^ TRIS base buffer, 0.0211 mol L^−1^ TRIS HCl buffer, 0.0735 mol L^−1^ lithium dodecyl sulfate, 1.08 mol L^−1^ glycerol, 0.5 mmol L^−1^ ethylenediaminetetraacetic acid, 0.1 mg·mL−1 PefaBloc SC (AEBSF) protease inhibitor) modified from Brown *et al*. [@brownFluxCapacitiesAcclimation2008]. Cells were lysed for 3 cycles of 60 seconds at 6.5m s-1 with 60 second intervals on ice between cycles.  The lysing vials were centrifuged at 16100 x *g* using a tabletop Mikro 20 centrifuge (Andreas Hettich GmbH & Co., Tuttlingen, Germany) for 5 minutes, the supernatant was transferred to a 600 µL microcentrifuge tube and centrifuged again at 16100 x *g* for 3 minutes. 40 µL of the supernatant were subsampled into labeled vials and stored at -80 °C until analysis. The bicinchoninic acid (BCA) assay was used for total protein determination using a calibration curve of bovine gamma globulin (BGG; Bio-Rad, Hercules, CA, USA) standard ranging from 0 to 1.0 mg mL^-1^ range.

@@

Using the dataset Omar *et al*. [@omarAnnotationGenesEncoding2023], we filtered the dataset for Enzyme Commission Numbers (EC numbers), or Kegg Orthology Numbers (KO numbers) identified by BRENDA  [@changBRENDAELIXIRCore2021] as 'natural substrates' for O~2~ in *P. marinus* (MED4, MIT9313, SS120, and NATL2A) in order to generate a subset containing only those orthologs encoding enzymes directly mediating O~2~ metabolism. We grouped orthologs together by EC number and their KO number and determined the occurrences of individual orthologs encoding each EC number, or KO number when EC number was not available, in a given strain. We merged the dataset with a list of enzyme Michaelis constant (K~m~) values from other organisms, as K~m~ values from *Prochlorococcus* were only available in the case of Ribulose bisphosphate carboxylase. XXX Possibly add the link to the full list of genes. Remove this once added XXXX

@@

NATLProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

Guided by the evidence of ocean distributions of proteins from *Prochlorococcus* we set up a matrix of photoperiods, PAR, spectral bands, and [O~2~] to approximate current, and potential future, latitudinal, depth and seasonal niches for *Prochlorococcus* strains.  We earlier found [@murphyPhotoinactivationPhotosystemII2017] that growth under red light could prove mechanistically informative as to factors limiting *Prochlorococcus* growth, so we included that spectral band even though it is not characteristic of representative *Prochlorococcus* niches.  Although *Prochlorococcus* is currently limited to a narrow range of surface photoperiods, potential poleward latitudinal expansions, in combination with attenuation of light with depth, mean *Prochlorococcus* may potentially encounter a wide range of photoperiods.  Our growth rate determinations generally agreed with those from @mooreComparativePhysiologySynechococcus1995, for white LED and 250 µM O~2~, but our study is, to our knowledge, the first to analyze the interactive growth responses of *Prochlorococcus* strains to varying [O~2~], spectral wavebands and photoperiods.

@@

Similar to growth trends under 250 µM O~2~, MED4 maintained at 25 µM O~2~ showed fastest growth when the
photoperiod was 16 h for each spectral waveband, across PAR levels (Figure \@ref(fig:MED4PAR)).
The µ~max~ was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 180 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod. The 4 h photoperiod experiments under white LED light were not performed as no growth was achieved when grown under an 8 h photoperiod of white LED light.

MED4 did not grow when sparged to the lowest [O~2~] of 2.5 µM (Figure \@ref(fig:MED4PAR)). 2.5 µM O~2~ growth experiments
were not conducted for 4 and 16 h photoperiods, as no reproducible growth occurred when MED4 was exposed to 8 and 12 h photoperiods.

@@

Under 250 µM O~2~ growth rates of *P. marinus* SS120, Clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white, red or blue LED light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

Under 25 µM O~2~ and PAR of 30 µmol photons m^-2^ s^-1^ growth trends were similar to 250 µM O~2~. SS120 showed no
growth under a 4 h photoperiod for red spectral waveband, however under blue light, SS120 was able to grow under a 4 h photoperiod (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew under 4 and 8 h blue light and 16 h red light photoperiods, however the growth rate decreased under 12 and 16 h white LED light photoperiod treatments. Blue light treatments under PAR of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h photoperiod. The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ blue light and 8 h photoperiod. The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^  under white LED light experiments were not performed due to time constraints.

@@

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments. Most 4 h photoperiod treatments of SS120 also did not grow under 250 µM O~2~, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments (Figure \@ref(fig:BluevsRedPurFitsPlots)B).
